---
title: "Quickstart guide for creating an R package on gitHub"
author: Lukas Dargel
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart guide for creating an R package on gitHub}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library("RgitDemo")
```


## Setup your Project on GitHub

### Quickstart

```{r, eval=FALSE}
# 0) optionally write default settings into your R profile
usethis::edit_r_profile()
usethis::edit_r_environ()

# 1) create this project
usethis::create_package("/Projects/Packages/RgitDemo")

# 2) create this vignette
usethis::use_vignette("Quickstart_R_package_on_gitHub", title = "Quickstart guide for creating an R package on gitHub")

# 3) initialize git
usethis::use_github()

# 4) create your first function
usethis::use_r("utils_list.R")

# 5) invite your collaborators to fork your repo and start working together.
```

### Some explanations


**Prerequisites** create an account on gitHub and download the RStudio IDE.

**In step 0** we may set some default values.

+ If you do not have already a GitHub token create it with `usethis::create_github_token()`.
+ Store the token inside RStudio by calling `gitcreds::gitcreds_set()` then pasting the token. 
This allows to use the gui to push and pull directly from the RStudio IDE.
+ 

**In step 1** we create a minimal R package based on RStudio projects


**In step 2** we create a template for an vignette based on Rmarkdown


**In step 3** we initialize git, start version control.


**In step 4** we create an R-script in the R directory and add a first function to our package.

**From Step 5 onwards** we start to develop the project collectively.
At this point it is recommended not to make any changes directly on the master/main branch.
Instead we use forks to develop new functionality and merge them into the main branch via pull requests, which are explained in the next section.

## Start collaborative development



